Review of Edge Detectors

### 2012-07-23 edge-detection-review-v4.tex
new file:   1dmasks_n.pdf 	new file:   2dconvolution.tex 	new file:   bibliography.bib 	new file:   blockdiagram1.pdf 	new file:   blockdiagram2.pdf 	new file:   edge-detection-review-v10.tex 	new file:   edge-detection-review-v2.tex 	new file:   edge-detection-review-v3.tex 	new file:   edge-detection-review-v4.tex 	new file:   edge-detection-review-v5.tex 	new file:   edge-detection-review-v6.tex 	new file:   edge-detection-review-v7.tex 	new file:   edge-detection-review-v8.tex 	new file:   edge-detection-review-v9.tex 	new file:   edge-detection-review.tex 	new file:   examples/lena.png 	n

\usepackage{hyperref,verbatim,graphicx,amsmath,amssymb,amssymb,dsfont}
\usepackage[ruled,linesnumbered]{algorithm2e}
\usepackage{sidecap,subfig,array}
\usepackage[table]{xcolor}
\usepackage{multicol}
\newtheorem{theorem}{Theorem}
\numberwithin{equation}{section}
\numberwithin{table}{section}
\numberwithin{figure}{section}
\begin{document}

@@

%-------------------------------------------------------------------------------
\begin{ipolAbstract}
In this paper we study some of the classic alternatives for edge detection in digital images. The main idea 
behind the edge detection is to find where there are abrupt changes in the intensity of an image. 
As a first approach, the first derivative is used to find the changes of intensity in several algorithms 
such as Sobel, Prewitt and Roberts. The second derivative is also used, for example in algorithms 
like Marr-Hildreth and Haralick.\\
We compare the results obtained from a qualitative point of view (perceptual) and from a quantitative 
point of view (number of operations, execution time), considering different forms to convolve an 
image with a kernel (step required in some of the algorithms).\\

@@

%-------------------------------------------------------------------------------
\begin{ipolCode}
The C implementation of the algorithms implemented in this work (beta version) can be
downloaded from:
\begin{itemize}
	\item Roberts, Prewitt and Sobel operators: \href{http://iie.fing.edu.uy/~haldos/ipol/fded.tar.gz}{here}.
	\item Marr-Hildreth algorithm: \href{http://iie.fing.edu.uy/~haldos/ipol/marr-hildreth.tar.gz}{here}.
	\item Haralick algorithm: \href{http://iie.fing.edu.uy/~haldos/ipol/haralick.tar.gz}{here}.
\end{itemize}
\end{ipolCode}

@@

Edge detection is one of the first approaches published in the literature for 
segmenting images. The basic idea is to detect abrupt changes in intensity. 
Detecting those changes in intensity for the purpose of finding edges in images 
can be accomplished using first or second order derivatives.

@@

In 1980 Marr and Hildreth argued that intensity changes are not independent 
of image scale, so the edge detection requires the use of different size 
operators. They also argued that a sudden intensity change will be seen 
as a peak (or trough) in the first derivative or, equivalently, as a zero 
crossing in the second derivative. The bulk of this work focuses on the 
study of this algorithm, presented in Section ~\ref{sec:second}.\\

In section ~\ref{sec:examples} we present some examples to compare the 
performance of the algorithms studied in this work. We also present a 
video created using the Marr-Hildreth algorithm, applied frame by frame.\\

@@

The selected tool to find the amplitude and direction of changes in 
intensity of an image $f$ is the gradient operator (denoted as $\nabla$), defined 
as the vector:
\begin{equation}
	\nabla f = 
				\begin{bmatrix} 
					g_x \\ g_y
				\end{bmatrix}
	=				
				\begin{bmatrix} 
					\cfrac{\partial f}{\partial x} \\ \cfrac{\partial f}{\partial y} \\
				\end{bmatrix}
\end{equation}

@@

We need to compute partial derivatives $\partial f/\partial x$ and $\partial f/\partial y$ 
at every pixel in the image. Since we are dealing with digital images, numerical approximations 
of the partial derivatives over a neighborhood about a point are required. Next we will study 
the methods of Roberts, Prewitt and Sobel, whose main difference is how they approach this calculation.\\

@@

The most usual approach of the first derivative is to take the Taylor expansion of order zero with
a small $h$. Approximating the first derivative:
\begin{equation}
	f'(x) \simeq \frac{f(x+h) - f(x)}{h}\\
\end{equation}
The image has discrete points of coordinates $(i,j)$, so taking $x=i*h$ and $y=j*h$, we approximate the 
first derivative of the image based on the intensity values ​​at points of the image as:
\begin{equation}
\label{eq:roberts1}
	g_x = \frac{\partial f(x,y)}{\partial x} = f(x+1,y) - f(x,y)
\end{equation}
and
\begin{equation}
\label{eq:roberts2}
	g_y = \frac{\partial f(x,y)}{\partial y} = f(x,y+1) - f(x,y)\\
\end{equation}

Equations \ref{eq:roberts1} and \ref{eq:roberts2} can be implemented for all values of $x$ and $y$
by filtering the image $f(x,y)$ with the 1-D masks in figure \ref{fig:1dmasks}. But when we want 
to find diagonal edges, we need 2-D masks. The \textit{Roberts operators} are one of the earliest 
attempts to use 2-D masks with this purpose. These operators are based on implementing the diagonal 
diferences implemented by filtering an image with the masks in figure \ref{fig:roberts}.\\

@@

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.2\textwidth]{results/square.png}
	\caption{Test image 1: white square on black background.}
	\label{fig:original1}
\end{figure}

The figures \ref{fig:result1-a} to \ref{fig:result1-f} show the output of each algorithm. The first three correspond to the first derivative methods (Roberts, Prewitt and Sobel), then the following two are from the Marr-Hildreth algorithm using Gaussian and LoG kernels, and the last one corresponds to the Haralick algorithm. Execution time\footnote{Running on Intel Core i3 CPU (2.53GHz), 3 Gb RAM, standard laptop.} for each algorithm are shown in Table \ref{exectime1}. \\

@@

It is appreciated that in the simplest cases, the first derivative edge detection algorithms are running better and faster. Let us now analyze the performance of each of the algorithmos using a more realistic test image, shown in figure \ref{fig:original2}. \\

\begin{figure}[t!]
	\centering
	\includegraphics[width=0.5\textwidth]{results/molino.jpg}
	\caption{Test image 2: windmill.}
	\label{fig:original2}
\end{figure}

First we show in figure \ref{fig:result2} the results of the first derivative edge detection algorithms.

\begin{figure}[h!]
	\centering
	\subfloat[Roberts.]{\label{fig:result2-a}\includegraphics[width=0.42\textwidth]{results/molino_roberts.png}}
	\quad
	\subfloat[Prewitt.]{\label{fig:result2-b}\includegraphics[width=0.42\textwidth]{results/molino_prewitt.png}}

	\subfloat[Sobel.]{\label{fig:result2-c}\includegraphics[width=0.42\textwidth]{results/molino_sobel.png}}
	\caption{Results of the first derivative algorithms using the figure \ref{fig:original2} as input image.}
	\label{fig:result2}
\end{figure}

Now we show in figures \ref{fig:result3-a} and \ref{fig:result3-b} the results obtained using the Marr-Hildreth algorithm (both Gaussian and LoG kernels), and in figure \ref{fig:result3-c} the results obtained using the Haralick algorithm. In table \ref{exectime2} we show the execution time for each one of the algorithms. \\

\begin{figure}[h!]
	\centering
	\subfloat[Marr-Hildreth (Gaussian).]{\label{fig:result3-a}\includegraphics[width=0.42\textwidth]{results/molino_marr-hildreth.png}}
	\quad
	\subfloat[Marr-Hildreth (LoG).]{\label{fig:result3-b}\includegraphics[width=0.42\textwidth]{results/molino_marr-hildreth-log.png}}

	\subfloat[Haralick.]{\label{fig:result3-c}\includegraphics[width=0.42\textwidth]{results/molino_haralick.png}}
	\caption{Results of the Marr-Hildreth and the Haralick algorithms, using the figure \ref{fig:original2} as input image.}
	\label{fig:result3}
\end{figure}

@@

It is now visible the difference in performance between the first derivative edge detection algorithms and the second derivative ones. \\

@@

\clearpage
%-------------------------------------------------------------------------------
\section{Appendix: Common operations}
\label{sec:appendix1}

@@

We generate the Gaussian kernel by sampling the 2-D Gaussian function\footnote{Note that for simplicity 
we omitted the normalizing coefficient $1/\sqrt{2\pi\sigma^2}$.}
\begin{equation}
	\label{eq:gaussian_function}
	G(x,y) = e^{-\frac{x^2+y^2}{2\sigma^2}}
\end{equation}
where $\sigma$ is the standard deviation (sometimes $\sigma$ is called the \textit{space constant}).\\

@@

Further examples obtained with the implemented algorithms are shown in figures \ref{fig:example1} to \ref{fig:example3}. More examples can be found in the online demo. \\

\begin{figure}[h!]
	\centering
	\subfloat[Roberts.]{\includegraphics[width=0.3\textwidth]{examples/lena.png}}

	\subfloat[Marr-Hildreth (Gaussian).]{\includegraphics[width=0.3\textwidth]{examples/lena_marr-hildreth.png}}
	\quad
	\subfloat[Marr-Hildreth (LoG).]{\includegraphics[width=0.3\textwidth]{examples/lena_marr-hildreth-log.png}}
	\quad
	\subfloat[Haralick.]{\includegraphics[width=0.3\textwidth]{examples/lena_haralick.png}}
	\caption{Example: Lena.}
	\label{fig:example1}
\end{figure}

\begin{figure}[t!]
	\centering
	\subfloat[Roberts.]{\includegraphics[width=0.3\textwidth]{examples/oranges.png}}

	\subfloat[Marr-Hildreth (Gaussian).]{\includegraphics[width=0.3\textwidth]{examples/oranges_marr-hildreth.png}}
	\quad
	\subfloat[Marr-Hildreth (LoG).]{\includegraphics[width=0.3\textwidth]{examples/oranges_marr-hildreth-log.png}}
	\quad
	\subfloat[Haralick.]{\includegraphics[width=0.3\textwidth]{examples/oranges_haralick.png}}
	\caption{Example: Oranges.}
	\label{fig:example2}
\end{figure}

\vspace{2cm}

\begin{figure}[t!]
	\centering
	\subfloat[Roberts.]{\includegraphics[width=0.3\textwidth]{examples/shapes.png}}

	\subfloat[Roberts.]{\includegraphics[width=0.3\textwidth]{examples/shapes_roberts.png}}
	\quad
	\subfloat[Prewitt.]{\includegraphics[width=0.3\textwidth]{examples/shapes_prewitt.png}}
	\quad
	\subfloat[Sobel.]{\includegraphics[width=0.3\textwidth]{examples/shapes_sobel.png}}

	\subfloat[Marr-Hildreth (Gaussian).]{\includegraphics[width=0.3\textwidth]{examples/shapes_marr-hildreth.png}}
	\quad
	\subfloat[Marr-Hildreth (LoG).]{\includegraphics[width=0.3\textwidth]{examples/shapes_marr-hildreth-log.png}}
	\quad
	\subfloat[Haralick.]{\includegraphics[width=0.3\textwidth]{examples/shapes_haralick.png}}
	\caption{Example: Shapes.}
	\label{fig:example3}
\end{figure}

%-------------------------------------------------------------------------------
\begin{thebibliography}{9}

\end{thebibliography}

\end{document}
%-------------------------------------------------------------------------------

>>>>

\usepackage{hyperref,verbatim,graphicx,amsmath,amssymb,amssymb,dsfont}
\usepackage[ruled,linesnumbered]{algorithm2e}
\usepackage{sidecap,subfig,array}
\usepackage[table]{xcolor}
\usepackage{multicol}
\usepackage{listings}
\usepackage{color}
\newtheorem{theorem}{Theorem}
\numberwithin{equation}{section}
\numberwithin{table}{section}
\numberwithin{figure}{section}
\begin{document}
\bibliographystyle{ieeetr}
\lstset{
	backgroundcolor=\color[rgb]{0.9,0.9,0.9},
    tabsize=4,
    inputencoding=utf8x,
         extendedchars=\true,
         language=C++,
         basicstyle=\scriptsize,
                 showtabs=false,
        showspaces=false,
        showstringspaces=false,
        identifierstyle=\ttfamily,
        keywordstyle=\color[rgb]{0,0,1},
        commentstyle=\color[rgb]{0.133,0.545,0.133},
        stringstyle=\color[rgb]{0.627,0.126,0.941},
}

@@

%-------------------------------------------------------------------------------
\begin{ipolAbstract}
In this paper we study some of the classic alternatives for edge detection in digital images. The main idea 
behind the edge detection is to find where there are abrupt changes in the intensity of an image. 
As a first approach, the first derivative is used to find the changes of intensity in several algorithms 
such as Sobel, Prewitt and Roberts \cite{Gonzalez2007Digital}. The second derivative is also used, for example in algorithms 
like Marr-Hildreth \cite{AIM-518} and Haralick \cite{bb20239}.\\
We compare the results obtained from a qualitative point of view (perceptual) and from a quantitative 
point of view (number of operations, execution time), considering different forms to convolve an 
image with a kernel (step required in some of the algorithms).\\

@@

%-------------------------------------------------------------------------------
\begin{ipolCode}
The C implementation of the algorithms implemented in this work (beta version) can be
downloaded from \href{http://iie.fing.edu.uy/~haldos/downloads/edge_detectors_v0.1.tar.gz}{here}.
%\begin{itemize}
%	\item Roberts, Prewitt and Sobel operators: \href{http://iie.fing.edu.uy/~haldos/ipol/fded.tar.gz}{here}.
%	\item Marr-Hildreth algorithm: \href{http://iie.fing.edu.uy/~haldos/ipol/marr-hildreth.tar.gz}{here}.
%	\item Haralick algorithm: \href{http://iie.fing.edu.uy/~haldos/ipol/haralick.tar.gz}{here}.
%\end{itemize}
% TODO referencia a github
\end{ipolCode}

@@

Edge detection is one of the first approaches published in the literature for 
segmenting images. The basic idea is to detect abrupt changes in intensity. 
Detecting those changes in intensity for the purpose of finding edges in images 
can be accomplished using first or second order derivatives. \\

@@

In 1980 Marr and Hildreth argued that intensity changes are not independent 
of image scale, so the edge detection requires the use of different size 
operators. They also argued that a sudden intensity change will be seen 
as a peak (or trough) in the first derivative or, equivalently, as a zero 
crossing in the second derivative. The bulk of this work focuses on the 
study of this algorithm, presented in Section ~\ref{sec:second}. In this section 
we also present the Haralick algorithm, which is also based on the second derivative.
This algorithm has the particularity of proposing a model to locally approximate 
the image around a point. Then, with this model, second derivative can be calculated 
analytically (rather approximate) and finding edges is done by imposing a 
condition over the model parameters. \\

Results and conclusions are shown in sections \ref{sec:results} and 
\ref{sec:conclusions} respectively. In section ~\ref{sec:examples} we present some examples to compare the 
performance of the algorithms studied in this work. We also present a 
video created using the implemented algorithms, applied frame by frame.\\

@@

The usual tool to find the amplitude and direction of changes in 
intensity of an image $f$ is the gradient operator (denoted as $\nabla$), defined 
as the vector:
\begin{equation}
	\nabla f = 
				\begin{bmatrix} 
					g_x \\ g_y
				\end{bmatrix}
	=				
				\begin{bmatrix} 
					\cfrac{\partial f}{\partial x} \\ \cfrac{\partial f}{\partial y} \\
				\end{bmatrix}
\end{equation}

@@

We need to compute partial derivatives $\partial f/\partial x$ and $\partial f/\partial y$ 
at every pixel in the image. Since we are dealing with digital images, numerical approximations 
of the partial derivatives over a neighborhood about a point are required. In the following we will study 
the methods of Roberts, Prewitt and Sobel, whose main difference is how they perform this calculation.\\

@@

The most usual approach of the first derivative is to take the Taylor expansion of first order with
a small $h$. Approximating the first derivative:
\begin{equation}
	f'(x) \simeq \frac{f(x+h) - f(x)}{h}\\
\end{equation}
The image has discrete points of coordinates $(i,j)$, so taking $x=i*h$ and $y=j*h$, we approximate the 
first derivative of the image based on the intensity values ​​at points of the image as:
\begin{equation}
\label{eq:roberts1}
	g_x = \frac{\partial f(x,y)}{\partial x} = f(i+1,j) - f(i,j)
\end{equation}
and
\begin{equation}
\label{eq:roberts2}
	g_y = \frac{\partial f(x,y)}{\partial y} = f(i,j+1) - f(i,j)\\
\end{equation}

Equations \ref{eq:roberts1} and \ref{eq:roberts2} can be implemented for all values of $x$ and $y$
by filtering the image $f(x,y)$ with the 1-D masks in figure \ref{fig:1dmasks}. But when we want 
to find diagonal edges, we need 2-D masks. The \textit{Roberts operators} are one of the earliest 
attempts to use 2-D masks for this purpose. These operators are based on implementing the diagonal 
diferences implemented by filtering an image with the masks in figure \ref{fig:roberts}.\\

@@

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.2\textwidth]{results/square.png}
	\caption{Test image 1: white square on black background (128$\times$128 pixels).}
	\label{fig:original1}
\end{figure}

Figures \ref{fig:result1-a} to \ref{fig:result1-f} show the output of each algorithm. The first three correspond to the first derivative methods (Roberts, Prewitt and Sobel), then the following two are from the Marr-Hildreth algorithm using Gaussian and LoG kernels, and the last one corresponds to the Haralick algorithm. Execution times\footnote{Running on Intel Core i3 CPU (2.53GHz), 3 Gb RAM, standard laptop.} for each algorithm are shown in Table \ref{exectime1}. \\

@@

It is appreciated that in the simplest cases, the first derivative edge detection algorithms are running better and faster. All algorithms show consistent results, but the Haralick algorithm shows some rare artifacts out of the edges. This may occur because this is the only algorithm that assumes a certain regularity (third degree) in the neighborhood of a pixel. Synthetic edges are far from this regularity. This may cause the adjustment of the bicubic polynomial function around these points to be quite different from reality. \\

Let us now analyze the performance of each of the algorithms using a more realistic test image, shown in figure \ref{fig:original2}. \\

\begin{figure}[t!]
	\centering
	\subfloat[Original image.]{\includegraphics[width=0.48\textwidth]{results/molino.jpg}}
	\quad
	\subfloat[Grayscale image.]{\includegraphics[width=0.48\textwidth]{results/molino_bw.png}}
	\caption{Test image 2: windmill (1000$\times$563 pixels).}
	\label{fig:original2}
\end{figure}

First we show in figures \ref{fig:result2-a} to \ref{fig:result2-c} the results of the first derivative edge detection algorithms. Figures \ref{fig:result2-d} and \ref{fig:result2-e} show the results obtained using the Marr-Hildreth algorithm (both Gaussian and LoG kernels), and figure \ref{fig:result2-f} shows the results obtained using the Haralick algorithm. In table \ref{exectime2} we show the execution time for each one of the algorithms. \\
\begin{figure}[h!]
	\centering
	\subfloat[Roberts.]{\label{fig:result2-a}\includegraphics[width=0.42\textwidth]{results/molino_roberts.png}}
	\quad
	\subfloat[Prewitt.]{\label{fig:result2-b}\includegraphics[width=0.42\textwidth]{results/molino_prewitt.png}}

	\subfloat[Sobel.]{\label{fig:result2-c}\includegraphics[width=0.42\textwidth]{results/molino_sobel.png}}
	\quad
	\subfloat[Marr-Hildreth (Gaussian).]{\label{fig:result2-d}\includegraphics[width=0.42\textwidth]{results/molino_marr-hildreth.png}}

	\subfloat[Marr-Hildreth (LoG).]{\label{fig:result2-e}\includegraphics[width=0.42\textwidth]{results/molino_marr-hildreth-log.png}}
	\quad
	\subfloat[Haralick.]{\label{fig:result2-f}\includegraphics[width=0.42\textwidth]{results/molino_haralick.png}}
	\caption{Results of the first derivative, Marr-Hildreth and Haralick algorithms using the figure \ref{fig:original2} as input image.}
	\label{fig:result2}
\end{figure}

@@

It is now visible the difference in performance between the first derivative edge detection algorithms and the second derivative ones. We can see in figure \ref{fig:result3} an area of ​​interest of the image, enlarged to better view of the detail. Note that none of the first derivative methods (even with a loose threshold) detect the lower edge of the blade (which is shaded). Neither Haralick algorithm detects it. However, the Marr-Hildreth algorithm detects it, using either a Gaussian or a LoG kernel. \\

Another observation is that the edges detected by the Haralick algorithm appear to be thicker than those detected with the other algorithms. This may be due the regularity that Haralick assumes, previously mentioned. \\

\begin{figure}[h!]
	\centering
	\subfloat[Original.]{\label{fig:result3-a}\includegraphics[width=0.48\textwidth]{results/molino_bw_zoom.png}}

	\subfloat[Roberts.]{\label{fig:result3-b}\includegraphics[width=0.48\textwidth]{results/molino_roberts_zoom.png}}
	\quad
	\subfloat[Prewitt.]{\label{fig:result3-c}\includegraphics[width=0.48\textwidth]{results/molino_prewitt_zoom.png}}

	\subfloat[Sobel.]{\label{fig:result3-d}\includegraphics[width=0.48\textwidth]{results/molino_sobel_zoom.png}}
	\quad
	\subfloat[Marr-Hildreth (Gaussian).]{\label{fig:result3-e}\includegraphics[width=0.48\textwidth]{results/molino_marr-hildreth_zoom.png}}

	\subfloat[Marr-Hildreth (LoG).]{\label{fig:result3-f}\includegraphics[width=0.48\textwidth]{results/molino_marr-hildreth-log_zoom.png}}
	\quad
	\subfloat[Haralick.]{\label{fig:result3-g}\includegraphics[width=0.48\textwidth]{results/molino_haralick_zoom.png}}
	\caption{Results of the algorithms using the figure \ref{fig:original2} as input image, enlarging an interesting area.}
	\label{fig:result3}
\end{figure}

@@

We have discussed and carefully implemented some of the more traditional methods of edge detection in digital images. We have tested the implemented algorithms with synthetic and real images, obtaining generally the expected results, taking into account the limitations of these methods. \\

The first derivative algorithms (Roberts, Prewitt and Sobel) have the advantage of having a very simple implementation. Also they run extremely fast, because they only consists of a convolution with a very small kernel (2$\times$2 or 3$\times$3 pixels). The results obtained with these methods are quite good, considering its simplicity, but they have problems such as noise and discontinuity at the edges. \\

The second derivative algoritms (Marr-Hildreth \& Haralick) involve several more operations, since more convolutions (and with larger kernels) are performed. In real images, they have better behavior towards first derivative algorithms. \\

Comparing the two versions of the Marr-Hildreth algorithm, is significantly faster the version with Gaussian kernel and numerical approximation of the Laplacian. The LoG kernel version, while slower (since it needs a larger kernel to achieve similar results) is more accurate, because it makes no approximation to calculate the Laplacian (it is calculated analytically before creating the kernel). It is also possible to manage the size of the kernel, which represents a scale parameter of the algorithm, being able to obtain a highly detailed edge image using small kernels, or just more noticeable edges using larger kernels. \\

Haralick's algorithm, although it shares the Marr-Hildreth idea of finding zero crossings of the second derivative, has some quite different ideas. It is the only one of these algorithms which works with an approximation of the intensity in the neighborhood of a pixel, using a bicubic polinimial function. This supposes a fairly large regularity in the image, which is not always true, and sometimes leads to detect edges where none of them exist, and get some thicker edges too. \\

All algorithms have their advantages and disadvantages. Choosing one or other may depend on requirements after edge detection. \\

% At the end:
Finally, this paper is a summary of some classical edge detection algorithms. In the preparation of this paper, no similar material (concentrated in one text) was found, so this paper appears to be useful material for whom starts studying edge detection methods. \\

Further examples obtained with the implemented algorithms are shown in figures \ref{fig:example1} and \ref{fig:example2}. More examples can be found in the online \href{http://link_to_demo}{demo}. \\

\begin{figure}[h!]
	\centering
	\subfloat[Original.]{\includegraphics[width=0.3\textwidth]{examples/lena.png}}
	\quad
	\subfloat[Grayscale.]{\includegraphics[width=0.3\textwidth]{examples/lena_bw.png}}

	\subfloat[Marr-Hildreth (Gaussian).]{\includegraphics[width=0.3\textwidth]{examples/lena_marr-hildreth.png}}
	\quad
	\subfloat[Marr-Hildreth (LoG).]{\includegraphics[width=0.3\textwidth]{examples/lena_marr-hildreth-log.png}}
	\quad
	\subfloat[Haralick.]{\includegraphics[width=0.3\textwidth]{examples/lena_haralick.png}}
	\caption{Example: Lena (512$\times$512 pixels).}
	\label{fig:example1}
\end{figure}

\begin{figure}[t!]
	\centering
	\subfloat[Original.]{\includegraphics[width=0.3\textwidth]{examples/oranges.png}}
	\quad
	\subfloat[Grayscale.]{\includegraphics[width=0.3\textwidth]{examples/oranges_bw.png}}

	\subfloat[Marr-Hildreth (Gaussian).]{\includegraphics[width=0.3\textwidth]{examples/oranges_marr-hildreth.png}}
	\quad
	\subfloat[Marr-Hildreth (LoG).]{\includegraphics[width=0.3\textwidth]{examples/oranges_marr-hildreth-log.png}}
	\quad
	\subfloat[Haralick.]{\includegraphics[width=0.3\textwidth]{examples/oranges_haralick.png}}
	\caption{Example: Oranges (536$\times$480 pixels).}
	\label{fig:example2}
\end{figure}

Some observations: 
\begin{itemize}
	\item The first derivative algorithms, although they work properly and quickly, have some problems like discontinuity in the edges. They are also very affected by noise in images, appearing unwanted isolated edges. \\
	\item The Marr-Hildreth algorithm (in its two versions) achieved interesting results in the edge detail. This algorithm provides a first step to regularity in the image, since the filtering with a Gaussian kernel. This behavior can be seen in Lena's hair (figure \ref{fig:example1}), or in the inside of the oranges (example \ref{fig:example2}). \\
	\item As mentioned earlier, Haralick's algorithm is the first to assume greater regularity of the image in the neighborhood of a pixel (third order). This causes thicker edges, as shown in both examples. But, note that some edges are only detected with this algorithm, e.g. the lower edges of the oranges, which are in the shadow, in figure \ref{fig:example2}. These edges are quite smooth, so that the Haralick model is well suited to them, and they do not represent an abrupt change in the intensity to be detected by other methods. \\
\end{itemize}

%\vspace{2cm}

%\begin{figure}[t!]
%	\centering
%	\subfloat[Original.]{\includegraphics[width=0.3\textwidth]{examples/shapes.png}}

%	\subfloat[Roberts.]{\includegraphics[width=0.3\textwidth]{examples/shapes_roberts.png}}
%	\quad
%	\subfloat[Prewitt.]{\includegraphics[width=0.3\textwidth]{examples/shapes_prewitt.png}}
%	\quad
%	\subfloat[Sobel.]{\includegraphics[width=0.3\textwidth]{examples/shapes_sobel.png}}

%	\subfloat[Marr-Hildreth (Gaussian).]{\includegraphics[width=0.3\textwidth]{examples/shapes_marr-hildreth.png}}
%	\quad
%	\subfloat[Marr-Hildreth (LoG).]{\includegraphics[width=0.3\textwidth]{examples/shapes_marr-hildreth-log.png}}
%	\quad
%	\subfloat[Haralick.]{\includegraphics[width=0.3\textwidth]{examples/shapes_haralick.png}}
%	\caption{Example: Shapes.}
%	\label{fig:example3}
%\end{figure}

%-------------------------------------------------------------------------------
%\begin{thebibliography}{9}
%\end{thebibliography}

\bibliography{bibliography}

\clearpage
%-------------------------------------------------------------------------------
\section{Appendix 1: Common operations}
\label{sec:appendix1}

@@

We generate the Gaussian kernel by sampling the 2-D Gaussian function\footnote{Note that for simplicity 
we omitted the normalizing coefficient $1/\sqrt{2\pi\sigma^2}$.}
\begin{equation}
	\label{eq:gaussian_function}
	G(x,y) = e^{-\frac{x^2+y^2}{2\sigma^2}}
\end{equation}
where $\sigma$ is the standard deviation (sometimes $\sigma$ is called the \textit{scale space constant}).\\

@@

\clearpage
%-------------------------------------------------------------------------------
\section{Appendix 2: Implementation}
\label{sec:appendix2}

\subsection{Roberts, Prewitt and Sobel}
\input{test_fded.tex}

\subsection{Marr-Hildreth}
\label{app:marr-hildreth}
\input{test_mh.tex}

\subsection{Haralick}
\input{test_haralick.tex}

\subsection{Gaussian kernel generation}
\input{gaussian_kernel.tex}

\subsection{2D convolution}
\input{2dconvolution.tex}

%-------------------------------------------------------------------------------
\end{document}
